---
title: "HW1"
author: "Zeqiu.Yu"
date: "2022-10-02"
output: pdf_document
---

Question 1.5  
No.
The simple linear model should be stated as 
\[Y_i = \beta_0 + \beta_1X_i + \epsilon_i\] or
\[E\{Y_i\} = \beta_0 + \beta_1X_i\]


Question 1.19
```{r, echo = TRUE, eval = TRUE, warning = TRUE, result = TRUE,include = TRUE}
input1 <- read.table("./CH01PR19.txt")
names(input1) <- c("GPA","ACT")
```
(a.) 
```{r, echo = TRUE, eval = TRUE, warning = TRUE, result = TRUE,include = TRUE}
fit1 <- lm(GPA~ACT, data = input1)
summary(fit1)
```
\[b_0 = 2.11405, b_1 = 0.03883.\] Then we get:
\[\hat{Y} = 2.11405 + 0.03883 X\]

(b.) Plot the estimated regression function and the data. Does the estimated regression function appear to fit the data well.
```{r, echo = FALSE, eval = TRUE, warning = FALSE, result = TRUE,include = TRUE}
require(ggplot2)
g <- ggplot(input1, aes(x = ACT, y = GPA))
g + geom_point(size = 2,color = "blue") + geom_smooth(method = lm, se = TRUE,linetype = "dashed",color = "red") + geom_rug()
```

The estimated regression funnction doesn't seem to fit it well. The distribustion of the points is so spread and the linear model has a relative low adjusted R-squared value 0.06476.

(c.) Obtain a point estimate of the mean fresh man GPA for students with ACT test score X = 30.
\[\hat{Y} = 2.11405 + 0.03883 X = 3.27895\]

(d) What is the point estimate of the change in the mean response when the entrance test score increase by one point?
\[1 \times b_1 = 0.03883\]

1.27
```{r, echo = TRUE, eval = TRUE, warning = TRUE, result = TRUE,include = TRUE}
input2 <- read.table("./CH01PR27.txt")
names(input2) <- c("muscle_mass","Age")
```
(a.) 
```{r, echo = TRUE, eval = TRUE, warning = TRUE, result = TRUE,include = TRUE}
fit2 <- lm(muscle_mass~Age, data = input2)
summary(fit2)
```

The esitimated regression function is:
\[\hat{Y} = 156.3466 -1.1900X\]
```{r, echo = FALSE, eval = TRUE, warning = FALSE, result = TRUE,include = TRUE}
require(ggplot2)
g <- ggplot(input2, aes(x = Age, y = muscle_mass))
g + geom_point(size = 2,color = "blue") + geom_smooth(method = lm, se = TRUE,linetype = "dashed",color = "red") + geom_rug()
```
The linear function appear to give good fit and the plot proves the anticipation that the muscle mass decreases with age.

(b.) 
(1) \[-1.1900\times 1 = -1.19\]
(2) \[\hat{Y} = 156.3466 -1.1900\times 60 = 84.9466\]
(3) For the eighth case, $X_8$ = 41 and $Y_8$ = 112
\[e_8 = Y_8 - \hat{Y_8} = 112 - (156.3466 -1.1900\times 41) = 4.4434\]
(4) We use MSE to estimate $\sigma^2$.
\[s^2 = \frac{1}{n-2}\sum_{i = 1}^n e_i^2\]
```{r, echo = TRUE, eval = TRUE, warning = TRUE, result = TRUE,include = TRUE}
squaredError <- (input2$muscle_mass - (156.3466 -1.1900*input2$Age))^2
s2 <- sum(squaredError)/(dim(input2)[1] - 2)
print(s2)
```
Hence, a point estimation of $\sigma^2$ is 66.80082.


(Remark: the values of $b_0$ and $b_1$ is consistent with the values calculated by the formula proved in the textbook.)

