---
title: "HW2"
author: "Zeqiu.Yu"
date: "2022-10-10"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Homework2  
## Question 2.27
```{r, echo = TRUE, eval = TRUE, warning = TRUE, result = TRUE,include = TRUE}
input1 <- read.table("./BookDataSets/Chapter  1 Data Sets/CH01PR27.txt")
names(input1) <- c("muscle_mass","Age")
```
a.  
By hypothesis test:
$H_0\;:\beta_1 = 0$ and $H_a\;:\beta_1 < 0$
```{r, echo = TRUE, eval = TRUE, warning = TRUE, result = TRUE,include = TRUE}
fit <- lm(muscle_mass~Age, data = input1)
summary(fit)
```
Then the test statistic $t^{*} = \frac{b_1 - 0}{s\{b_1\}} = \frac{(-1.19) - 0}{0.0902} = -13.1929$  
$H_0\;:\beta_0 \geq 0$ and $H_a\;:\beta_0 < 0$
```{r, echo = TRUE, eval = TRUE, warning = TRUE, result = TRUE,include = TRUE}
t_value = -qt(0.95, 58)
p_value = pt(-13.1929,58, lower.tail = TRUE)
print(sprintf("The t-value is:%.4f, and the p-value is: %.2f", t_value, p_value))
```
We accept $H_a$ because $t^{*}<t-value$

b. The answer is yes. The p-value is so small(<0.05) that can be used to reject $H_0\; : \beta_0 = 0$. That is to say, $b_0$ provides information.  
c. 
```{r, echo = TRUE, eval = TRUE, warning = TRUE, result = TRUE,include = TRUE}
confint(fit, "Age", level=0.95)
```
$\hat{Y_i} = b_0 + b_1X$ and $\hat{Y_{i+1}} = b_0 + b_1(X+1)$, $\Delta\hat{Y_i} = \hat{Y_{i+1}}-\hat{Y_i} = b_1$. Hence, it is not necessary to know the specific ages to make this estimate.

## Question 2.28  
a. 
```{r, echo = TRUE, eval = TRUE, warning = TRUE, result = TRUE,include = TRUE}
alpha = 0.05
x_new = data.frame(Age = 60)
predict(fit, x_new, interval = "confidence", level = 1-alpha, se.fit = TRUE)
```
The confidence interval is [82.83471, 87.05895]. We are 95% confident that the mean muscle mass for women aged 60 is between 82.83471 to 87.05895.  

b.  
```{r, echo = TRUE, eval = TRUE, warning = TRUE, result = TRUE,include = TRUE}
alpha = 0.05
x_new = data.frame(Age = 60)
predict(fit, x_new, interval = "prediction", level = 1-alpha)
```
The prediction interval is [68.45067, 101.443]. It's not in comparision with the result of a.  

c.  
```{r, echo = TRUE, eval = TRUE, warning = TRUE, result = TRUE,include = TRUE}
alpha = 0.05
x_new = data.frame(Age = 60)
CI <- predict(fit, x_new, interval = "prediction", level = 1-alpha, se.fit = TRUE)
Y_h_hat <- CI$fit[1]
Y_h_hat.se <- CI$se.fit
W <- sqrt(2*qf(1-alpha, 2, 58))
Band <- c(Y_h_hat - W*Y_h_hat.se, Y_h_hat + W*Y_h_hat.se)
Band
```
Yes. It is wider and it should be, because the confidence band encompass the entire regression line and one is able to draw conclusion about any values of X.  

## Question 2.29
a.  
```{r, echo = FALSE, eval = TRUE, warning = FALSE, result = TRUE,include = TRUE}
library(ggplot2)
# library(gridExtra)
library(cowplot)
DF <- data.frame(input1$Age, input1$muscle_mass, -1.19*input1$Age + 156.3466)
names(DF) <- c("X","Y","Y_hat")
p1 <- ggplot(DF, aes(x = X, y = Y-Y_hat)) + 
  geom_point(size = 2,color = "blue") + 
  theme_classic() +
  theme(plot.background = element_rect(color = "black"))
p2 <- ggplot(DF, aes(x = X, y = Y_hat-mean(Y))) + 
  geom_point(size = 2,color = "blue") + 
  theme_classic() +
  theme(plot.background = element_rect(color = "black"))
plot_grid(p1, p2, ncol = 2, align = "v")
```
$\hat{Y_i} - \bar{Y}$ has larger range and and then SSR has larger component. $R^2 = \frac{SSR}{SSTO}$, $R^2$ tend to be larger and therefore the proportion of the variation of Y can be "explained" by X the regression on X is larger.  

b. 
```{r, echo = TRUE, eval = TRUE, warning = TRUE, result = TRUE,include = TRUE}
anova(fit)  
```
c. Let $H_0\; : \beta_1 = 0$ and $H_a\; : \beta_1 \neq 0$.
$f^{*} \leq$ the critial value, I fail to reject $H_0$, otherwise accept $H_a$.  
$f^{*} = \frac{11627.5/1}{66.8} = 174.0623>F(1-\alpha; 1, 58) = 4.006873$. Then we accept $H_a$.  
d.
```{r, echo = TRUE, eval = TRUE, warning = TRUE, result = TRUE,include = TRUE}
VA = anova(fit)
SSR <- VA$"Sum Sq"[1]
SSE <- VA$"Sum Sq"[2]
R_sq <- SSR/(SSR+SSE)
1- R_sq
```
e. 
$R^2 = 0.7500668$, $r = -0.866064$

