---
title: "HW7"
author: "Zeqiu.Yu"
date: "2022-11-22"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Homework 7
## 14.7  
```{r}
Data <- read.table("./BookDataSets/Chapter 14 Data Sets/CH14PR07.txt",header=FALSE)
colnames(Data) <- c("Y", "X")
```

(a.) 
```{r}
glm.out = glm(Y~X, family=binomial(logit), data=Data)
summary(glm.out)
```

The maximum likelihood estimates: $\beta_0=-4.80751$ and $\beta_1=0.12508$, and  
$\hat{\pi  }=\frac{exp(-4.80751+0.12508X)}{1+exp(-4.80751+0.12508X)}$.  

(b.)
```{r}
plot(Y~X, data=Data)
lines(Data$X[order(Data$X)], glm.out$fitted[order(Data$X)], 
       type="l", col="red")
title(main="Fitted Logistic Regression Line")

Data.smooth <- predict(loess(Y~X, data=Data, span=0.75))
points(Data$X[order(Data$X)], Data.smooth[order(Data$X)], 
      type="b",lty=2, col="green")
legend(5,0.9, c("logistic","loess smooth"), col=c("red", "green"), lty=c(1:2))
```
The fitted logistic response function fitts appears to fit well.

## c.
```{r}
exp_b1 <- exp(glm.out$coefficients[2])
exp_b1
```

exp(b1) is `r exp_b1`. For an one-unit increase in the dollar increase in annual dues, we expect to see about 13.33% increase in the odds of that the membership will not be renewed.

## d.
```{r}
new <- data.frame(X=40, Y=1)
#y.hat <- predict(glm.out, new)      #predict log(p/(1-p))
#p.hat <- exp(y.hat)/(1+exp(y.hat))  #Transfer to probability
# or use option of "response" directly
p.hat <-predict(glm.out, new, type="response") 
p.hat
```

The estimated probability that association members will not renew their membership
if the dues are increased by $40 is `r p.hat`.

## e.
```{r}
(log(0.75/0.25)+4.80751)/0.12508
```

Since $\hat{\pi}$ = 0.75, $log(\frac{\hat{\pi}}{1-\hat{\pi}})=log(\frac{0.75}{0.25})=-4.80751+0.12508X$. Hence, X = 47.2187583.  

## 14.15  
(a.)
```{r}
conf.beta1 <- confint(glm.out, "X", level = 0.90)
exp(conf.beta1)
```
An approximate 90 percent confidence interval for $exp(\beta_1)$ is [1.021468, 1.276781].  
With 90 percent confidennce, we believe that, for an one-unit increase in the dollar increase in annual dues, the odds of that the membership will not be renewed will increase by between 2.1468% and 27.6781%.

## b)
```{r}
summary(glm.out)
qnorm(0.95, lower.tail=FALSE)
```

$H_0$: $\beta_1=0$   
$H_a$: $\beta_1\neq0$.$\\$  
Decision rule: $z^*=\frac{b_1}{s\{b_1\}}=\frac{0.12508}{0.06676}=1.873577$, $z(1-0.1/2)=z(0.95)=1.645$, if $|z*|\geq z(0.95)$, conclude $Ha$.  
Conclusion: $|z*|=1.873577 \ge z(0.95)=1.645$. Hence, conclude $H_a$.
